---
title: "Ch 21 Notebook"
output: html_notebook
---

```{r setup}
library(tidyverse)
```

##21.2 Exercises

###Exercise 1
  
Write for loops to:
  
  1. Compute the mean of every column in "mtcars"
```{r}
mtcars %>% map_dbl(mean)
```
  
  2. Determine the type of each column in "nycflights13::flights"
```{r}
yo<-nycflights13::flights

output <-vector("list", ncol(yo))
for (i in seq_along(yo)){
  output[[i]]<-class(yo[[i]])
}
output
```
  
  3. Compute the number of unique values in each column of "iris"
```{r}
output<-vector("list", ncol(iris))
for(i in seq_along(iris)){
  output[[i]]<-length(unique(iris[[i]]))
}
output
```
  
  
  4. Generate 10 random normals for each of M = -10, 0, 10, and 100
```{r}
#WHAT DOES THAT IT EVEN MEAN
```

###Exercise 2

Eliminate the for loop in each of the following examples by taking advantage of an existing function that works with vectors
```{r}
#Uhhhhhh
```


###Exercise 3

Combine your function writing and for loop skills:
  1. Write a for loop that "prints()" the lyrics to the children's song "Alice the camel"
```{r}
humps<-c("five","four","three","two","one","no")

alice<- function(x){
        cat("Alice the camel has", x ,"humps\nAlice the camel has", x, "humps\nAlice the camel has", x, "humps\nSo go, Alice, go\n\n")
}

for(i in seq_along(humps)){
  alice(humps[[i]])
}
```
  
  2. Convert the nursery ryhme "ten in the bed" to a function. Generalise it to any number of people in any sleeping structure.
```{r}
bed<- function(x){
  cat("There were", x, "in the bed\nAnd the little one said, \"Roll over! Roll over!\"\nSo they all rolled over and one fell out\n\n")
}

last<-"There was 1 in the bed\nAnd the little one said,\n\"Alone at last!\""

number<-function(x){
  y=x
  if(x>1){
    for(i in 2:x){
      bed(y)
      y=y-1
    } 
  } 
  cat(last)
}
number(1)
```
  
  
  3. Convert the song "99 bottes of beer on the wall" to a function. Generalise to any number of any vessel containing any liquid on any surface
```{r}
beer<-function(x,y){
  cat(x, "bottles of", y, "on the wall,", x, "bottles of",str_c( y, ".\nTake one down and pass it around,"), x-1, "bottles of", y, "on the wall\n\n")
}

yup<-function(x,y){
  z=x
  if(z>1){
    for(i in 2:x){
      beer(z,y)
      z=z-1
    } 
  }
  cat("No more bottles of", y, "on the wall, no more bottles of", y,".\nGo to the store and buy some more,",x,"bottles of",y,"on the wall")
}

yup(4,"milk")
  
```

###Exercise 4

It's common to see for loops that don't preallocate the output and instead increase the length of a vector at each step:
```{r}
x<-c(1,2,4,8,7,9)

output <- vector("integer", 0)
for (i in seq_along(x)){
  output <- c(output, length(x[[i]]))
}
  output 
```


##21.3 Exercises

###Exercise 1
Imagine you have a directory full of CSV files that you want to read in. You have their paths in a vector, "files <- dir("data/", pattern = "\\.csv$", full.names = TRUE)", and now want to read each one with read_csv(). Write the for loop that will load them into a single data frame.
```{r}
files <- dir(path="data/", pattern = "\\.csv$", full.names = TRUE)

output<-vector("list", length(files))
for (i in seq_along(files)){
  output[[i]]<-read.csv(files[[i]])
output
  }
```


###Exercise 2
What happens if you use "for (nm in names(x))" and x has no names? What if only some of the elements are named? What if the names are not unique?

```{r}
#No names:
x<-c(1, 2, 3)
for(nm in names(x)){
  print(nm) 
}
```
The environment tells me that nm is NULL! I don't
know why its showing up in the console, though?

```{r}
#Some named:
y<-c(do=1, re=2, mi=3, 4, so=5,la=6, ti=7)
for(nm in names(y)){
  print(nm)
}
```
Ahh, the value missing a name is given a placeholder name: ""

```{r}
#Same name:
z<-c(do=1, re=2, mi=3, la=4, so=5,la=6, ti=7, do=8)
for(nm in names(z)){
  print(nm)
}
```
"do" and "do" are considered different values.

###Exercise 3
Write a function that prints the mean of each numeric column in a data frame, along with its name. For example, "show_mean(iris)" would print:
```{r}
okay<-function(x){
  for (i in seq_along(x)){
    if(class(x[[i]])=="numeric"){
     cat(str_c(colnames(x[i]),":"),mean(x[,i], na.rm=FALSE),"\n") 
    }
    else{
      cat(str_c(colnames(x[i]),": NA"))
    }
  }
}
okay(iris)
```

###Exercise 4
What does this code do? How does it work?
```{r}
trans <- list( 
  disp = function(x) x * 0.0163871,
  am = function(x) {
    factor(x, labels = c("auto", "manual"))
  }
)
for (var in names(trans)) {
  mtcars[[var]] <- trans[[var]](mtcars[[var]])
}
```

##21.4 Exercises

###Exercise 1
Read the documentation for "apply()". In the 2d case, what two for loops does it generalise?
```{r}
?apply
```


###Exercise 2
Adapt "col_summary()" so that it only applies to numeric columns You might want to start with an "is_numeric()" function that returns a logical vector that has a TRUE corresponding to each numeric column.
```{r}
col_summary <- function(df, fun) {
  out <- vector("double", length(df))
  for (i in seq_along(df)) {
    if(is.numeric(df[[i]])==TRUE)
      out[i] <- fun(df[[i]])
  }
  out
}

col_summary(iris, mean)
```


##21.5 Exercises

###Exercise 1
Write code that uses one of the map functions to:
  
  1. Compute the mean of every column in "mtcars"
```{r}
mtcars %>% map_dbl(mean)
```
  
  2. Determine the type of each column in "nycflights13::flights"
```{r}
nycflights13::flights %>% map(class)
```
  
  3. Compute the number of unique values in each column of "iris"
```{r}
iris %>% map_int(~length(unique(.)))
```
  
  4. Generate 10 random normals for each of M = -10, 0, 10, and 100
```{r}
#I still don't get this at all :D
```

###Exercise 2
How can you create a single vector that for each column in a data frame indicates whether or not it’s a factor?
```{r}
(x<-iris %>% map_lgl(~is.factor(.)))
```

###Exercise 3
What happens when you use the map functions on vectors that aren’t lists? What does "map(1:5, runif)" do? Why?
```{r}
blah<-c(1,2,3,4,5,6)
map_dbl(blah, ~.*2)
```
Hmm, it seems fine? No problems here!

```{r}
map(1:5, runif)

map(2:5, runif)
```
Okay, so Google tells me "runif" is random number generator using Uniform Distribution...something, something...look, that's not important! What is impoprtant is that map(x:y) creates a list of y-x rows. The first row contained x random numbers; the number of random numbers increases by 1 per row until it reaches y random numbers. Err...did you follow all that? 

###Exercise 4
What does "map(-2:2, rnorm, n = 5)" do? Why? What does "map_dbl(-2:2, rnorm, n = 5)" do? Why?
```{r}
map(-2:2, rnorm, n = 5)

map_dbl(-2:2, rnorm, n = 5)
```
Okay, so quick Google search tells me rnorm(n, mean, sd) of normally distributed numbers. n is the sample size produced, mean is the sample mean (defaults to 0), and sd is teh sample standard deviation (defaults to 1). In this case, -2:2 represents five rows with means -2,-1,0,1,2, respectively. n=5 is sample size for each row.

map_dbl(-2:2, rnorm, n=5) produces an error. rnorm produces a set of numbers which is definitely NOT a doube vector.

##21.9 Exercises

###Exercise 1
Implement your own version of "every()" using a for loop. Compare it with "purrr::every()". What does purrr’s version do that your version doesn’t?
```{r}
ooh<-c(1,2,3,4,5)
aah<-c(1,2,"ack",4)

haha<-function(x,y){
  for(i in seq_along(x)){
    if(class(x[i]) != y){
      print(FALSE)
      break
    }
    if(i==length(x-1))
      print(TRUE)
  }
}

haha(ooh, "numeric")
haha(aah, "numeric")
```


###Exercise 2
Create an enhanced "col_summary()" that applies a summary function to every numeric column in a data frame.
```{r}
col_summary <- function(df) {
  for (i in seq_along(df)) {
    if(is.numeric(df[[i]])==TRUE)
      cat(c(colnames(df[i]),":\n"))
      print(summary(df[[i]]))
      cat("\n")
  }
}

col_summary(iris)
```